---
title: "NBA Shot Defense Analysis"
author: "Aidan Judge"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(ggmap)
library(viridis)
library(forcats)
library(smodels)
library(smodels, lib.loc = "C:/Program Files/R/R-3.4.0/library")
```

```{r}
nba_shots <- read_csv("https://statsmaths.github.io/stat_data/nba_shots.csv")
```
##Thesis

The average distance of the closest defender on shots will have a negative relationship with the shooting team's margin of victory for the 2014-2015 NBA season, even in close games decided by eight points or less.

##Exposition for Models 1 & 2:

Defense is a significant factor in any team's success in the NBA, and players like Kawhi Leonard and Lebron James are praised as being dynamic "two-way" players.  Sometimes, however, defensive effects can be challenging to quantify.  The average distance of the closest defender on shots will serve as a proxy for the quality of defense contesting the shots in the graphic below.  In the two graphs below, I have graphed the relationship between the average distance of the closest defender with the margin of victory first for all games in the 2015 NBA season and then again for the subset of "close games" decided by 8 points or fewer.  Both of the summary tables are displayed and show that there is a significant positive relationship between these variables on a 95% confidence interval.  For the dataset of all games, if the defenders give the shooters an extra foot of space on average, the margin of victory increases by an estimated 5.851 points.  For the dataset of close games decided by 8 points or fewer, if the defenders give the shooters an extra foot of space on average, the margin of victory increases by an estimated 1.4207 points.  This means that for every increase in the average distance of the closest defender, the margin of victory increased for the offensive team that was shooting.

#Model 1:

```{r}
individualgames <- group_summarize(nba_shots, matchup)
qplot(close_def_dist_mean, final_margin_mean, data = individualgames) +
  geom_smooth(method = "lm") +
  xlab("Average Distance of the Closest Defender") +
  ylab("Final Margin of Victory/Loss for Shooting Team") +
  ggtitle("Relationship of Defender Distance and NBA Game Outcomes")
model1 <- lm_basic(final_margin_mean ~ 1 + close_def_dist_mean, data = individualgames)
reg_table(model1, level = .95)
```

#Model 2:

```{r}
closegames <- filter(individualgames, final_margin_mean >= -8, final_margin_mean <= 8)
qplot(close_def_dist_mean, final_margin_mean, data = closegames) +
  geom_smooth(method = "lm") +
  xlab("Average Distance of the Closest Defender") +
  ylab("Final Margin of Victory/Loss for Shooting Team") +
  ggtitle("Relationship of Defender Distance and NBA Outcomes in Close Games")
model2 <- lm_basic(final_margin_mean ~ 1 + close_def_dist_mean, data = closegames)
reg_table(model2, level = .95)
```

##Exposition for Models 3 & 4:

With that being said, it is worthwhile to consider whether this statistic is always a good proxy for the quality of defense.  In the third graph below, I use a subset of the top 100 scorers in the 2015 season to show some of the most "unguardable" players in the NBA.  What we find is that the shooters with the highest shooting percentages for the most part are the tallest players, power forwards and centers, who spend the least time with the ball in their hands meaning they shoot quickly and give even physically close defenders no time to position themselves.  These players are also the ones who move around the least, and thus, as the fourth graphic shows, actually have defenders guarding them physically the closest.  The combination of these two graphs would challenge the original assumption that the average distance of the closest defender from the shooter is a strong proxy for the quality of defense.  While that might be true for the shorter guards playing the 1, 2, or 3 position, it is more challenging to quantify defensive quality in close quarters for the big men around the rim.  This does not lead me to challenge the hypothesis, but is food for thought on how to best defend different sized players in different positions.  Perhaps it could be a good starting point on different defensive formations: box-and-one vs. man-to-man vs. zone defense.  I still believe that guarding your opponents closely (i.e. keeping you distance low) will increase your chances of victory, just as making the extra pass on offense to find your most open teammate will increase your chances of victory.

```{r}
byplayers <- group_summarize(nba_shots, fct_lump(player_name, 100))
byplayers$player_name <- byplayers$'fct_lump(player_name, 100)'
qplot(shooter_height_mean, fgm_mean, data = byplayers) +
  geom_smooth(method = "lm") +
  geom_point(aes(color = cut(touch_time_mean, 5))) +
  xlab("Shooter Height") +
  ylab("Percentage of Field Goals Made") +
  ggtitle("Relationship of Shooters' Heights and FG% for Top 100 Scorers") +
  labs(color = "Average Time Shooter Holds the Ball")
qplot(close_def_dist_mean, shooter_height_mean, data = byplayers) +
  geom_smooth(method = "lm") +
  geom_point(aes(color = cut(touch_time_mean, 5))) +
  xlab("Average Distance of the Closest Defender") +
  ylab("Shooter Height") +
  ggtitle("Relationship of Shooters' Heights and How Closely They Are Guarded") +
  labs(color = "Average Time Shooter Holds the Ball")
```




