---
title: "Bronx Analysis"
author: "Aidan Judge"
output: html_notebook
---

**BELOW ARE INSTRUCTIONS FOR HOW TO COMPLETE THE ASSIGNMENT;
MAKE SURE THAT YOU DELETE ALL OF THE COMMENTS IN ALL CAPS
(AS WELL AS THIS COMMENT), BEFORE SUBMITTING**

THIS FIRST CHUNK SHOULD LOAD ALL OF THE PACKAGES (I THINK
YOU WILL NOT NEED TO CHANGE THIS)

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
library(readr)
library(ggplot2)
library(dplyr)
library(ggmap)
library(viridis)
library(forcats)
```

##Loading large dataset and filtered New York metropolitan dataset
```{r}
tracts <- read_csv("https://raw.githubusercontent.com/statsmaths/stat_data/gh-pages/acs_tract_data_2016.csv.gz")

nymet <- filter(tracts, cbsa == "New York-Newark-Jersey City, NY-NJ-PA")
```

##Creation of new variables
```{r}
nymet$leave_before_0700 <- nymet$leave_0000_0459 + nymet$leave_0500_0529 + nymet$leave_0530_0559 + nymet$leave_0600_0629 + nymet$leave_0629_0659

tracts$leave_before_0700 <- tracts$leave_0000_0459 + tracts$leave_0500_0529 + tracts$leave_0530_0559 + tracts$leave_0600_0629 + tracts$leave_0629_0659

temp <- select(nymet, starts_with("race_"))
nymet$max_race_category <- names(temp)[apply(temp, 1, which.max)]
```

##What parts of the NY Metropolitan area have large black populations
```{r}
qmplot(lon, lat, data = nymet, geom = "blank") +
  geom_point(aes(color = race_black, size = population), alpha = 0.3) +
  scale_color_viridis() +
  ggtitle("NY Metropolitan Area by Black Populations") +
  labs(size = "Population", color = "Percentage of Residents Identifying as Black")
```

##What parts of the NY Metropolitan area have large percentages of residents who were born in the state they now live in
```{r}
qmplot(lon, lat, data = nymet, geom = "blank") +
  geom_point(aes(color = born_same_st, size = population), alpha = 0.3) +
  scale_color_viridis() +
  ggtitle("NY Metropolitan Area by Black Populations") +
  labs(size = "Population", color = "Percentage of Residents Identifying as Black")
```


```{r}
qmplot(lon, lat, data = nymet, geom = "blank") +
  geom_point(aes(color = race_black, size = population), alpha = 0.3) +
  scale_color_viridis()


ggplot(nymet, aes(population, hh_income_q5)) +
  geom_point()

qmplot(lon, lat, data = nymet, geom = "blank") +
  geom_point(aes(color = leave_before_0700), alpha = 0.8) +
  scale_color_viridis()

ggplot(nymet, aes(population, hh_income_q5)) +
  geom_point(aes(color = max_race_category))
```

```{r}  



```

NOW, YOUR REPORT SHOULD GO HERE. FEEL FREE TO SPLIT IT UP
WITH SECTIONS `##`, BUT THIS IS NOT REQUIRED. ADD CODE CHUNKS
TO PRODUCE GRAPHICS WHERE THEY BELONG IN THE OUTPUT

## References

IF YOU CITE ANY EXTERNAL SOURCES INCLUDE PARANTHETICAL
REFERENCES [SUCH AS: (Arnold, 2017)] IN THE TEXT AND
ADD CITATIONS HERE.


