---
title: "NBA Analysis"
author: "Aidan Judge"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(ggmap)
library(viridis)
library(forcats)
library(smodels)
library(smodels, lib.loc = "C:/Program Files/R/R-3.4.0/library")
```

AVAILABLE DATASETS (choose one and delete the others)

```{r}
nba_shots <- read_csv("https://statsmaths.github.io/stat_data/nba_shots.csv")
```

##Model 1 Relationship between average defender distance on margin of victory/loss

```{r}
individualgames <- group_summarize(nba_shots, matchup)
qplot(close_def_dist_mean, final_margin_mean, data = individualgames) +
  geom_smooth(method = "lm")
model1 <- lm_basic(final_margin_mean ~ 1 + close_def_dist_mean, data = individualgames)
reg_table(model1, level = .95)
```

##Model 2 - same as 1 but with median defender distance

```{r}
qplot(close_def_dist_median, final_margin_mean, data = individualgames) +
  geom_smooth(method = "lm")
model2 <- lm_basic(final_margin_mean ~ 1 + close_def_dist_median, data = individualgames)
reg_table(model2, level = .95)
```

##Model 3 - Same as Model 1, but is it more significant for closer games (decided by 10 or less)

```{r}
closegames <- filter(individualgames, final_margin_mean >= -10, final_margin_mean <= 10)
qplot(close_def_dist_mean, final_margin_mean, data = closegames) +
  geom_smooth(method = "lm")
model3 <- lm_basic(final_margin_mean ~ 1 + close_def_dist_mean, data = closegames)
reg_table(model3, level = .95)
```

##Model 4 - Same as Model 2, but is it more significant for closer games (decided by 10 or less)

```{r}
qplot(close_def_dist_median, final_margin_mean, data = closegames) +
  geom_smooth(method = "lm")
model4 <- lm_basic(final_margin_mean ~ 1 + close_def_dist_median, data = closegames)
reg_table(model4, level = .95)
```

##Model 5 Group summarize on top 30 players and see if total shot volume relates to consistency, i.e. shooting percentage
```{r}
top30players <- group_summarize(nba_shots, fct_lump(player_name, 30))
top30players$player_name <- top30players$`fct_lump(player_name, 30)`
qplot(close_def_dist_mean, fgm_mean, data = top30players) +
  geom_smooth(method = "lm") +
  geom_text(aes(label = player_name), size = 2) +
  geom_point(aes(color = cut(shooter_height_mean, 5)))
             
model5 <- lm_basic(fgm_mean ~ 1 + close_def_dist_mean + shooter_height_mean, data = top30players)
reg_table(model5, level = .95)
```


##Thesis (Accompanies Model 4):
Defense is a significant factor in a team's success in the NBA, and players like Kawhi Leonard and Lebron James are praised as being dynamic "two-way" players, but defensive effects can be challenging to quantify.  The average distance of the closest defender on shots will have a negative relationship with the shooting team's margin of victory, particularly in close games decided by ten points or less.